Community biomass and species richness temporal trends reveal contrasting assembly and disassembly processes in natural food-webs

### 2022-05-02 paper/manuscript.Rmd
semeff results + results + discussion

---
title: "Community biomass and species richness temporal trends reveal contrasting assembly and disassembly processes in natural food-webs"
author: Alain Danet, Loubna El Madouri, Maud Mouchet, Elisa Thébault and Colin Fontaine
date: \today
output:
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: true
  bookdown::word_document2:
fontsize: 12pt
header-includes:
   - \usepackage{natbib}
   - \usepackage{float}
   - \usepackage{multirow}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
number_sections: true
---

@@

```{r}
loadd(rsq_table_bm_rich_mono_stable_trends)
loadd(reg_table_bm_rich_mono_stable_trends)
```

@@

---
## Variations of community biomass and species richness are predicted to affect community structure
#Changes of species richness and community biomass are indeed
#predicted to affect community structure. The positive link between biodiversity
#and ecosystem functioning in experimental studies predict that loss and
#gain of species richness should respectively result in loss and gain of
#community biomass [@tilman_biodiversity_2014; @pennekamp_biodiversity_2018]. But
#this relation is not always found in natural settings [@grace_integrative_2016;
#@li_relationship_2018; @maureaud_biodiversityecosystem_2019].
#
#Variations of species richness can also result in variation of evenness, i.e.
#equitability of species abundances or biomasses. Meta-analysis reported that
#higher evenness of abundances is linked to higher plant biomass in grasslands
#and primary productivity in marine ecosystems [@soininen_relationship_2012].
#Furthermore, studies reported that higher evenness plays a role of insurance in
#face of perturbations [@wittebolle_initial_2009]. Indeed, most studies reported
#positive relationship between species richness and evenness, but also neutral
#and negative ones [@soininen_relationship_2012; @ma_species_2005;
#@maureaud_biodiversityecosystem_2019].
## Variations of community biomass and species richness are predicted to affect food-web structure
---

@@

## Relationships among temporal trends

We modelled the relationships among the temporal trends of food-web structure
($dN$), that is connectance, average trophic level and evenness of biomass, and
the temporal trends of community biomass ($dB$) and species richness ($dR$) with
Ordinary Least Squares. The simultaneous assessment of the
effects of species richness and community biomass allows to tease apart their
respective effect on food-web structure.

$dN_j = \alpha + \beta_1dB_j + \beta_2dR_j + \epsilon_j$

With:

  - $dN$: temporal trends of food-web structure: connectance, average trophic
    level and evenness of biomass across species
  - $dB$: temporal trend of ($\log_e$) community biomass
  - $dR$: temporal trend of ($\log_e$) species richness
  - $j$: the site
  - $\alpha$: intercept 
  - $\beta$: slope 
  - $\epsilon$: gaussian error

In order to control for the uncertainties of the estimated temporal trends, each
temporal trend was weighted by the inverse of the product of the standard errors
of the estimated trends, i.e. estimated slopes for the response and the
explicative variables ($w_j = \dfrac{1}{s_{r_j} s_{e,j}}$, $w_j$ being the
weight, $s_{r_j}$ and $s_{e_j}$ being respectively the standard errors on the
temporal trends of the response and explicative variables). We also included a
statistical model including the interaction between the temporal trends of
species richness and community biomass (i.e.  adding $\beta_3dB_jdR_j$), the
results are presented in supplementary material (Fig.  SXX). To model the
relationships between the temporal trends of community biomass and the one of
species richness, we built the model as $dB_j = \alpha_j + \beta_1dR_j +
\epsilon_j$.

## Relationships along spatial gradient

Spatial variation of food-web structure are partly driven by environmental gradients, so
we controlled relationships among variables by adding two PCA axis
synthesizing respectively (1) stream gradient from upstream to downstream and
(2) enrichment (Fig. \@ref(fig:matmet)). Enrichment was characterized by the
average Biological Oxygen Demand [BOD, @mallin_factors_2006][^2]. We further
added a random effect of the hydrographic basin on the intercept.

[^2]: commentaire de colin: il faudrait introduire spatial gradient des l'intro
et cette partie est plutot sur les relationships among spatial gradient of biodiv facets

$N_j = \alpha + \beta_1B_j + \beta_2R_j + \beta_3E_{1,j} + \beta_4E_{2,j} + \epsilon_j + \epsilon_{\alpha|i}$

With:

  - $N_j$: median food-web structure 
  - $B_j$: median community biomass ($g.m^-2$) 
  - $R_j$: median species richness ($m^-2$) 
  - $E_{1,j}$: First PCA axis 
  - $E_{2,j}$: Second PCA axis
  - $\epsilon_{\alpha|i}$: random effect of the hydrographic basin
    $i$ on $\alpha$

To model the relationships between community biomass and species richness along
the spatial gradient, we built the equation as $B_j = \alpha + \beta_1R_j +
\beta_3E_{1,j} + \beta_4E_{2,j} + \epsilon_j$.

In order to compare the magnitude of the effects of community biomass and
species richness in both temporal and spatial models, we computed standardized
slopes ($r_{\delta}$) and their associated standard errors as $r_{\delta} =
\beta\frac{s_x}{s_y}$.  With $\beta$ being the unstandardised slope, $s_x$ and
$s_y$ the standard deviation of the explicative and response variables.

@@

```{r}
rsq_piel_nind <- rsq_table_bm_rich_mono_stable_trends %>%
  filter(Response == "piel_nind") %>%
  .$R.squared %>%
  round(., 3)

rsq_tlvl <- rsq_table_bm_rich_mono_stable_trends %>%
  filter(Response == "w_trph_lvl_avg") %>%
  .$R.squared %>%
  round(., 3)

rsq_tlvl_sp <- sp_rsq_table_bm_rich_mono_stable_trends %>%
  filter(Response == "w_trph_lvl_avg") %>%
  .$Conditional %>%
  round(., 3)

rsq_piel_bm_sp <- sp_rsq_table_bm_rich_mono_stable_trends %>%
  filter(Response == "piel_bm") %>%
  .$Conditional %>%
  round(., 3)
```
Over the sampling sites in the dataset, we found that
`r res_trends[res_trends$variable == "log_bm_std", ]$total[1]` stations displayed monotonic trends for biomass. Among them, 
`r res_trends[res_trends$variable == "log_bm_std" & res_trends$direction == "stable", ]$percent`%
were classified as stable,
`r res_trends[res_trends$variable == "log_bm_std" & res_trends$direction == "decrease", ]$percent`%
as decreasing,
`r res_trends[res_trends$variable == "log_bm_std" & res_trends$direction == "increase", ]$percent`%
as increasing. Regarding the species richness, we found that
`r res_trends[res_trends$variable == "log_rich_std", ]$total[1]` stations displayed monotonic trends.
`r res_trends[res_trends$variable == "log_rich_std" & res_trends$direction == "stable", ]$percent`%
were classified as stable,  
`r res_trends[res_trends$variable == "log_rich_std" & res_trends$direction == "decrease", ]$percent`%
as decreasing, and 
`r res_trends[res_trends$variable == "log_rich_std" & res_trends$direction == "increase", ]$percent`% as increasing.

Overall, we found high heterogeneity in the explicative power of statistical
models linking temporal dynamics of food-web structure, evenness of biomass  and the one of community
biomass and species richness spanning from $R_{adj}^2= `r rsq_piel_nind`$ for
the evenness of abundances to $R_{adj}^2= `r rsq_tlvl`$ for average trophic
level (Table SXX). The models over the spatial gradient were more performant, with
$R_{cond}^2$ varing from $`r rsq_tlvl_sp`$ to $`r rsq_piel_bm_sp`$ (Table
SXX), $R_{cond}^2$ being the $R^2$conditional to the random effect.

@@

p_bef <- fig1_2$rich$log_bm_std +
  annotate("text", x = .1, y = -.5, parse = T,
    label = beta_rsq(
      rsq = coeff_bef["rsq"],
      beta = coeff_bef["estimate"],
      se = coeff_bef["std.error"]
    )
   ) +
  labs(
    y = lab_log_bm,
    x = lab_log_rich
  ) +
scale_xy_log_trends(breaks = c(-.10, -.050, 0, .050, .10), axis = "x") +
scale_xy_log_trends(breaks = c(-.4, -.2, 0, .4, .2), axis = "y", sep = " / ")

We found a strong positive relationship between the temporal trends of community
biomass and the ones of species richness ($r_\delta = `r tps_r_sest["p_bm"]`$,
Fig. \@ref(fig:bef)), meaning that both temporal trends
had the same direction in average. At the sites 
where species richness decreased over time, community biomass also decreased
over time. On the other hand, at the sites where species richness increased over
time, community biomass also increased over time. Along the spatial gradient, we also
found a positive relationship between median community biomass and median
species richness, but half the strength as the temporal one
($r_\delta = `r sp_b_sest["rich"]`$).

The temporal trends of connectance were positively linked with the ones of
community biomass ($r_\delta = `r tps_bm_sest["ct"]`$), also meaning that both
temporal trends had the same direction in average. That is, the sites where
food-web connectance decreased over time also had community biomass that
decreased over time and reversely. It suggests that communities that decreased
in biomass over time loose generalist predators over time, as connectance
measures the degree of generalism, and reversely that communities
gaining biomass over time also gain generalist predators. Spatially, we also found
a positive relationship between median connectance and median species
richness ($r_\delta = `r sp_b_sest["ct"]`$). However, we did not observe
significant relationships between the temporal trends of connectance and the
ones of species richness (Fig. \@ref(fig:bmr)D) while we found a stong
negative relationship between median connectance and median species richness
along the spatial gradient
($r_\delta = `r sp_r_sest["ct"]`$, Fig. \@ref(fig:std)).

@@

Evenness of biomass was negatively linked to community
biomass in time and in space 
($r_\delta = `r tps_bm_sest["p_bm"]`$ and $`r sp_b_sest["p_bm"]`$ resp.; Fig.
\@ref(fig:bmr)C, Fig. \@ref(fig:std)). Regarding time, it means that food-webs that
lost biomass over time also had a more even distribution of biomass across species over
time and *vice versa*. A similar reasoning applies in space. On the
other hand, we found a positive relationship between the temporal trends of
biomass evenness and the one of species richness (Fig. \@ref(fig:bmr)F,
$r_\delta = `r tps_r_sest["p_bm"]`$). We found the same relationship in space,
twice as strong however ($r_\delta = `r sp_r_sest["p_bm"]`$).

@@

```{r}
fig2 <- fig1_2
fig2$rich$log_bm_std <- NULL
fig2$rich <- map(fig2$rich,
  ~.x +
    theme(
      axis.text = element_text(size = 6),
      axis.title = element_text(size = 8)
      )
  )
fig2$bm <- map2(fig2$bm, names(fig2$rich),
  ~ .x + theme(
    axis.text = element_text(size = 6),
    axis.title = element_text(size = 8)
    ) +
  labs(
    y = paste0(var_replacement()[.y],
      "\ntrends (per year)"),
    x = lab_log_bm
  ) +
  scale_xy_log_trends(breaks = c(-.40, -.20, 0, .20, .40), axis = "x")
)
# Delete redondant legend
fig2$bm[[1]] <- fig2$bm[[1]] + theme(
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.x = element_blank()
  )

@@

```{r}
# Add beta and Rsq
for (i in seq_along(fig2$rich)) {

@@

```{r sem, fig.cap = sem_cap}
ti <- magick::image_read(here::here("figures", "sem_tps_avg_weight_ok.svg"),  density = 300)
ggdraw() +
  draw_image(ti,
    x = 0, y = 0, hjust = 0, vjust = 0, width = .9, height = 1)
```

```{r bmr, out.width="100%", fig.height=7, fig.cap=fig_bm_rich_cap}
p <- draw_image(here::here("figures", "sem_tps_avg_weight_ok.pdf"),
  x = 1, y = 1, hjust = 1, vjust = 1, width = 0.13, height = 0.2)
biplot_model <- plot_grid(
  plot_grid(, p_bef_hist,
    labels = c("", "A"), ncol = 2,
    rel_widths = c(.62, 1)
    ),
  plot_grid(
  plot_grid(plotlist = fig2$bm,
    ncol = 1, labels = c("B", "D"),
    hjust = -5.5, vjust = 1.75
    ),
  plot_grid(plotlist = fig2$rich,
    ncol = 1, labels = c("C", "E"),
    hjust = -2, vjust = 1.75
    ),
  ncol = 2,
  rel_widths = c(.85, 1)
  ),
  nrow = 2,
  rel_heights = c(.6, 1)
)
save_plot(
  filename = here::here("figures", "biplot_model.png"),
  biplot_model)

biplot_model
```

@@

```{r std, fig.cap=std_cap, fig.width=6, fig.height = 9}
load(here::here("data", "semeff.rda"))
load(here::here("data", "sp_semeff.rda"))
sp_semeff_tab <- from_semEff_to_table(sp_semeff)
semeff_tab <- from_semEff_to_table(semeff)

semeff_tot <- rbind(
  tibble(semeff_tab, model = "temporal"),
  tibble(sp_semeff_tab, model = "spatial")
  ) %>%
filter(
  !predictor %in% c("log_RC1", "log_RC2"),
  effect_type != "mediators"
)

semeff_tot %>%
  mutate(
    response = var_replacement()[response],
    predictor = var_replacement()[predictor]
    ) %>%
  ggplot(aes(x = predictor, y = effect, color = model)) +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(ymin = lower_ci, ymax = upper_ci), position = position_dodge(width = 0.4)) +
  facet_grid(rows = vars(response), cols = vars(effect_type)) +
  labs(
    x = "Predictor",
    y = expression(paste("Standardized slope coefficients ", r[delta])),
    color = "Model") +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

@@

We found that most temporal trends of food-webs were classified as
stable for species richness and community biomass. Roughly the same
number of stations displayed temporal decrease and increase for species
richness, and twice more stations displayed decrease in community biomass than
increase. This result is in accordance with previous meta analysis on species
richness trends of vertebrate in marine, freshwater and terrestrial systems
[@dornelas_assemblage_2014; @blowes_geography_2019; @klink_meta-analysis_2020].

@@

Spatially and temporally, species richness scaled positively with community
biomass. Spatially, this result is in accordance with previous studies in stream
communities [@woods_testing_2020; @danet_species_nodate] but also in grasslands,
salt marches and forests [@hector_plant_1999; @tilman_biodiversity_2014;
@wu_relationship_2015; @li_relationship_2018; but see @grace_integrative_2016].
Therefore, our results confirm that Biodiversity-Ecosystem functioning
prediction works also in the temporal dynamics of food-webs. The positive
relationship between species richness and total biomass might be due to positive
complementarity in resource use [@carey_determining_2011], which is held by
niche differentiation [@loreau_biodiversity_1998; @loreau_partitioning_2001] or
to the higher probability of getting high biomass species with higher species
richness [@loreau_biodiversity_2001].

@@

While we found a strong relationship between the temporal trends of community
biomass and the one of species richness, our results suggest that temporal
trends in community biomass and species richness are associated with very
different temporal trends on the connectance, average trophic level and evenness
of biomass. Indeed, temporal trends of connectance and average trophic level
were positively associated to the temporal trends of community biomass while no
significant association was found with temporal trends of species richness.
Moreover, the temporal trends of biomass evenness were negatively associated to
the temporal trends of community biomass whereas positively associated to
temporal trends of species richness. All together, it suggests that temporal
trends of community biomass and species richness are related to different
assembly and disassembly processes in food-webs.

The temporal trends of connectance were positively linked to that of community
biomass, meaning that when connectance increased through time when community
biomass increased through time in a given site. Reversely, connectance decreased
in food-webs when community biomass decreased through time.  Indeed, food-webs
have a highly nested structure, in a way that feeding trophic links of species
at lower trophic level are a subset of the ones of species found at higher
trophic level [@kondoh_food_2010; @nordstrom_nestedness_2015]. It is because
feeding interactions are largely driven by body size, consumers being larger
than their preys [@brose_allometric_2006; @woodward_body_2005]. We also found
that community biomass losses through time were associated to a temporal
decrease of average trophic level and that community biomass temporal gains were
associated to temporal increase of average trophic level. Together with connectance, it suggests that
community biomass losses and gains are associated to losses and gains in large
bodied piscivores, i.e. biomass at the higher trophic level. Another piece of evidence
is that evenness of biomass, measured by the Pielou index, decreased through
time as community biomass increased, meaning that biomass distribution across
species becomes less even when community biomass increased. Reversely, in
food-webs that displayed temporal decrease in community biomass, we found that
evenness of biomass increased, i.e. that biomass distribution across species
becomes more even. Because predator-prey interactions relate on bodysize for a
large part [@brose_allometric_2006; @woodward_body_2005] and thus bodymass to a
certain extent, we can expect that a given gain in predators of higher trophic
level leads to a decrease of biomass evenness and reversely. In the same line,
we found that the temporal trends of community biomass were positively
associated to the temporal trends of the number and proportion of piscivorous
trophic species present in fish communities (Fig. SXX). The close association
between the temporal trends of community biomass from one side, and connectance,
average trophic level and biomass evenness from the other side, strongly suggest
that temporal trends in community biomass in natural freshwater fish communities
are associated to variations in the biomass of the top trophic level.

>>>>

---
title: "Community biomass and species richness temporal trends reveal contrasting assembly and disassembly processes in natural food-webs"
author: Alain Danet, Loubna El Madouri, Maud Mouchet, Elisa Thébault and Colin Fontaine
date: \today
output:
  bookdown::word_document2:
    fontsize: 12pt
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: true
    fontsize: 12pt
    header-includes:
       - \usepackage{natbib}
       - \usepackage{float}
       - \usepackage{multirow}
       - \usepackage{booktabs}
       - \usepackage{makecell}
       - \usepackage{setspace}
       - \doublespacing
       - \usepackage{lineno}
       - \linenumbers
    geometry: margin=2cm
    number_sections: true
bibliography: references.bib
---

@@

```{r}
loadd(rsq_table_bm_rich_mono_stable_trends)
loadd(reg_table_bm_rich_mono_stable_trends)
load(here::here("data", "semeff.rda"))
load(here::here("data", "sp_semeff.rda"))
```

@@

---
## Variations of community biomass and species richness are predicted to affect
community structure #Changes of species richness and community biomass are
indeed
#predicted to affect community structure. The positive link between biodiversity
#and ecosystem functioning in experimental studies predict that loss and
#gain of species richness should respectively result in loss and gain of
#community biomass [@tilman_biodiversity_2014; @pennekamp_biodiversity_2018]. But
#this relation is not always found in natural settings [@grace_integrative_2016;
#@li_relationship_2018; @maureaud_biodiversityecosystem_2019].
#
#Variations of species richness can also result in variation of evenness, i.e.
#equitability of species abundances or biomasses. Meta-analysis reported that
#higher evenness of abundances is linked to higher plant biomass in grasslands
#and primary productivity in marine ecosystems [@soininen_relationship_2012].
#Furthermore, studies reported that higher evenness plays a role of insurance in
#face of perturbations [@wittebolle_initial_2009]. Indeed, most studies reported
#positive relationship between species richness and evenness, but also neutral
#and negative ones [@soininen_relationship_2012; @ma_species_2005;
#@maureaud_biodiversityecosystem_2019].
## Variations of community biomass and species richness are predicted to affect food-web structure

---

@@

We modelled the relationships among food-web structure, that is
connectance and average trophic level, and community biomass ($dB$) and species
richness ($dR$) with Structural Equation Models. We modelled a it such as species
richness is related to community biomass and food-web structure. In
turn community biomass was related to food-web structure. We computed one
structural equation model to assess the relationships among the temporal trends
and one for the relationships over the spatial gradient. As spatial variation of
food-web structure are partly driven by environmental gradients, so we
controlled relationships among variables by adding two PCA axis
synthesizing respectively (1) stream gradient from upstream to downstream and
(2) enrichment (Fig. \@ref(fig:matmet)). Enrichment was characterized by the
average Biological Oxygen Demand [BOD, @mallin_factors_2006][^2]. We further
added a random effect of the hydrographic basin on the intercept.

In order to control for the uncertainties of the estimated temporal trends, each
temporal trend was weighted by the inverse of the sum of the standard errors
of the estimated trends, i.e. estimated slopes for the response and the
explicative variables ($w_j = \dfrac{1}{\sum{e_j}}$, $w_j$ being the
weight, and $s_{e_j}$ the standard error on the
temporal trends of the variables).

The coefficients of the Structural Equation Models were computed with
PiecewiseSEM R package [@lefcheck_piecewisesem_2016]. The coefficients were
standardizated and the confidence intervals were computed with semEff R package
[@murphy_semEff_2021]. Confidence intervals were obtained by bootstrapping the
linear models 1000 times and computing adjusted bootstrap percentile interval
(BCa), which are adjusted for asymetric distribution of resampling.

@@

```{r}
rsq_ct <- rsq_table_bm_rich_mono_stable_trends %>%
  filter(Response == "ct_ff") %>%
  .$R.squared %>%
  round(., 2)

rsq_bm <- rsq_table_bm_rich_mono_stable_trends %>%
  filter(Response == "log_bm_std") %>%
  .$R.squared %>%
  round(., 2)

rsq_tlvl_sp <- sp_rsq_table_bm_rich_mono_stable_trends %>%
  filter(Response == "w_trph_lvl_avg") %>%
  .$Conditional %>%
  round(., 2)

rsq_ct_sp <- sp_rsq_table_bm_rich_mono_stable_trends %>%
  filter(Response == "ct_ff") %>%
  .$Conditional %>%
  round(., 2)
```
Over the sampling sites in the dataset, we found that
`r res_trends[res_trends$variable == "log_bm_std", ]$total[1]` stations displayed monotonic trends for biomass. Among them, 
`r res_trends[res_trends$variable == "log_bm_std" & res_trends$direction == "stable", ]$percent`%
were classified as stable,
`r res_trends[res_trends$variable == "log_bm_std" & res_trends$direction == "decrease", ]$percent`%
as decreasing,
`r res_trends[res_trends$variable == "log_bm_std" & res_trends$direction == "increase", ]$percent`%
as increasing. Regarding the species richness, we found that
`r res_trends[res_trends$variable == "log_rich_std", ]$total[1]` stations displayed monotonic trends.
`r res_trends[res_trends$variable == "log_rich_std" & res_trends$direction == "stable", ]$percent`%
were classified as stable,  
`r res_trends[res_trends$variable == "log_rich_std" & res_trends$direction == "decrease", ]$percent`%
as decreasing, and 
`r res_trends[res_trends$variable == "log_rich_std" & res_trends$direction == "increase", ]$percent`% as increasing.

Overall, we found relatively low in the explicative power of statistical
models linking temporal dynamics of food-web structure and the one of community
biomass and species richness spanning from $R_{adj}^2= `r rsq_ct`$ for
connectance to $R_{adj}^2= `r rsq_bm`$ for community biomass (Table SXX).
The models over the spatial gradient were more performant, with
$R_{cond}^2$ varing from $`r rsq_tlvl_sp`$ for the average trophic level to
$`r rsq_ct_sp`$ for connectance (Table SXX), $R_{cond}^2$ being the
$R^2$conditional to the random effect.

@@

We found a strong positive relationship between the temporal trends of community
biomass and the ones of species richness ($r_\delta = `r tps_r_sest["p_bm"]`$,
Fig. \@ref(fig:bmr) B), meaning that both temporal trends
had the same direction in average. At the sites 
where species richness decreased over time, community biomass also decreased
over time. On the other hand, at the sites where species richness increased over
time, community biomass also increased over time. Along the spatial gradient, we also
found a positive relationship between median community biomass and median
species richness, but half the strength as the temporal one
($r_\delta = `r sp_b_sest["rich"]`$).

The temporal trends of connectance were positively linked with the ones of
community biomass ($r_\delta = `r tps_bm_sest["ct"]`$), also meaning that both
temporal trends had the same direction in average. That is, the sites where
food-web connectance decreased over time also had community biomass that
decreased over time and reversely. It suggests that communities that decreased
in biomass over time loose generalist predators over time, as connectance
measures the degree of generalism, and reversely that communities
gaining biomass over time also gain generalist predators. Spatially, we also found
a positive relationship between median connectance and median species
richness ($r_\delta = `r sp_b_sest["ct"]`$). We found a negative 
relationship between the temporal trends of connectance and the
ones of species richness ($r_\delta = `r tps_r_sest["ct"]`$, Fig. \@ref(fig:bmr)D)
while we found negative relationship three times stronger between median connectance and
median species richness along the spatial gradient 
($r_\delta = `r sp_r_sest["ct"]`$, Fig. \@ref(fig:std)).

@@

```{r}
sp_semeff_tab <- from_semEff_to_table(sp_semeff)
semeff_tab <- from_semEff_to_table(semeff)
semeff
semeff_tot <- rbind(
  tibble(semeff_tab, model = "temporal"),
  tibble(sp_semeff_tab, model = "spatial")
  ) %>%
filter(
  !predictor %in% c("log_RC1", "log_RC2"),
  effect_type != "mediators"
)
```

```{r}
semeff_tab_coeff <- semeff_tab %>%
  mutate(across(where(is.numeric), ~round(., 2))) %>%
  mutate(eci = paste0(effect, " [", lower_ci, ";", upper_ci, "]"))

sp_semeff_tab_coeff <- sp_semeff_tab %>%
  mutate(across(where(is.numeric), ~round(., 2))) %>%
  mutate(eci = paste0(effect, " [", lower_ci, ";", upper_ci, "]")) %>%
  filter(! predictor %in% c("log_RC1", "log_RC2"))

ctsi <- semeff_tab_coeff %>%
  filter(response == "ct_ff", effect_type == "indirect")
ctstr <- semeff_tab_coeff %>%
  filter(response == "ct_ff", effect_type == "total",
    predictor == "log_rich_std")

ctspi <- sp_semeff_tab_coeff %>%
  filter(response == "ct_ff", effect_type == "indirect")
ctsptr <- sp_semeff_tab_coeff %>%
  filter(response == "ct_ff", effect_type == "total",
    predictor == "log_rich_std")

wtsi <- semeff_tab_coeff %>%
  filter(response == "w_trph_lvl_avg", effect_type == "indirect")
wtstr <- semeff_tab_coeff %>%
  filter(response == "w_trph_lvl_avg", effect_type == "total",
    predictor == "log_rich_std")

wtspi <- sp_semeff_tab_coeff %>%
  filter(response == "w_trph_lvl_avg", effect_type == "indirect")
wtsptr <- sp_semeff_tab_coeff %>%
  filter(response == "w_trph_lvl_avg", effect_type == "total",
    predictor == "log_rich_std")
```

The structural model on the temporal trends of food-web structure reveals that
the species richness has a direct negative effect on connectance but also a
negative one, twice lower than the former, through community biomass
($r_\delta$ = `r ctsi[["eci"]]`). It resulted a relatively weak total negative
effect of species richness on connectance (`r ctstr[["eci"]]`). Spatially,
we also found a positive indirect effect of richness on connectance through
community biomass but a weak one (`r ctspi[["eci"]]`), resulting in a strong negative total
effect of richness on connectance (`r ctsptr[["eci"]]`).

The relationships among temporal trends showed a indirect positive effect of
species richness on average trophic level (`r wtsi[["eci"]]`), resulting in a total positive
effect of species richness on average trophic level (`r wtstr[["eci"]]`). Over the spatial
gradient, we found the same positive indirect effect of richness through
community biomass (`r wtspi[["eci"]]`). But as the direct effect of richness on connectance
was strongly negative, it resulted in a total strong negative effect of species
richness on connectance over the spatial gradient (`r wtsptr[["eci"]]`).

@@

p_bef <- fig1_2$rich$log_bm_std +
  labs(
    y = lab_log_bm,
    x = lab_log_rich
  ) +
scale_xy_log_trends(breaks = c(-.10, -.050, 0, .050, .10), axis = "x") +
scale_xy_log_trends(breaks = c(-.4, -.2, 0, .4, .2), axis = "y", sep = " / ") +
    theme(
      axis.text = element_text(size = 8),
      axis.title = element_text(size = 10)
      )

```{r}
fig2 <- fig1_2
fig2$rich$log_bm_std <- NULL
fig2$rich <- map(fig2$rich,
  ~.x +
    theme(
      axis.text = element_text(size = 8),
      axis.title = element_text(size = 10)
      )
  )
fig2$bm <- map2(fig2$bm, names(fig2$rich),
  ~ .x + theme(
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 10)
    ) +
  labs(
    y = paste0(var_replacement()[.y],
      "\ntrends (per year)"),
    x = lab_log_bm
  ) +
  scale_xy_log_trends(breaks = c(-.40, -.20, 0, .20, .40), axis = "x")
)
# Delete redondant legend
fig2$bm[[1]] <- fig2$bm[[1]] + theme(
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.x = element_blank()
  )

@@

```{r, eval = FALSE}
# Add beta and Rsq
for (i in seq_along(fig2$rich)) {

@@

```{r sem, fig.cap = sem_cap}
ti <- magick::image_read(here::here("figures", "sem_tps_avg_weight_ok.svg"),  density = 300)
p <- ggdraw() +
  draw_image(ti,
    x = 0, y = 0, hjust = 0, vjust = 0, width = .9, height = 1)
```

```{r bmr, out.width="100%", fig.height=7, fig.cap=fig_bm_rich_cap}
#p <- draw_image(here::here("figures", "sem_tps_avg_weight_ok.pdf"),
#  x = 1, y = 1, hjust = 1, vjust = 1, width = 0.13, height = 0.2)
biplot_model <- plot_grid(
  plot_grid(NULL, p_bef_hist,
    labels = c("", "A"), ncol = 2,
    rel_widths = c(.54, 1)
    ),
  plot_grid(
  plot_grid(plotlist = fig2$bm,
    ncol = 1, labels = c("B", "D"),
    hjust = -5.5, vjust = 1.75,
    rel_heights = c(.85, 1)
    ),
  plot_grid(plotlist = fig2$rich,
    ncol = 1, labels = c("C", "E"),
    hjust = -2, vjust = 1.75,
    rel_heights = c(.85, 1)
    ),
  ncol = 2,
  rel_widths = c(.9, 1)
  ),
  nrow = 2,
  rel_heights = c(.55, 1)
)
sem_p <- plot_grid(p, biplot_model, ncol = 2, rel_widths = c(.4, .6))

save_plot(
  filename = here::here("paper", "figures", "biplot_tps.png"),
  biplot_model, nrow = 3, ncol = 2, base_asp = 1)
save_plot(
  filename = here::here("paper", "figures", "sem_p.png"),
  sem_p, nrow = 3, ncol =3)
sem_p
```

@@

```{r std, fig.cap=std_cap, fig.width=6, fig.height = 9}
semeff_tot %>%
  mutate(
    response = var_replacement()[response],
    predictor = var_replacement()[predictor]
    ) %>%
  ggplot(aes(x = predictor, y = effect, fill = effect_type)) +
  geom_col(
    position = position_dodge(width = 0.4), width = .4) +
  geom_hline(yintercept = 0) +
  geom_linerange(
    aes(ymin = lower_ci, ymax = upper_ci),
    position = position_dodge(width = 0.4)
    ) +
  facet_grid(rows = vars(response), cols = vars(model)) +
  labs(
    x = "Predictor",
    y = expression(paste("Standardized slope coefficients ", r[delta])),
    fill = "Effect type") +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

@@

Most temporal trends of for species richness and community biomasss were
classified as stable. Roughly the same number of stations displayed temporal
decrease and increase for species richness, and twice more stations displayed
decrease in community biomass than increase. This result is in accordance with
previous meta analysis on species richness trends of vertebrate in marine,
freshwater and terrestrial systems, which show no change in species richness in
average [@dornelas_assemblage_2014; @blowes_geography_2019] but an increase in
total abundance [@klink_meta-analysis_2020].

@@

Spatially and temporally, species richness scaled positively with community
biomass. Spatially, this result is in accordance with previous studies in stream
communities [@woods_testing_2020; @danet_species_nodate] but also in grasslands,
salt marches and forests [@hector_plant_1999; @tilman_biodiversity_2014;
@wu_relationship_2015; @li_relationship_2018; but see @grace_integrative_2016].
Therefore, our results confirm that Biodiversity-Ecosystem functioning
prediction also hold in the temporal dynamics of food-webs. The positive
relationship between species richness and total biomass might be due to positive
complementarity in resource use [@carey_determining_2011], which is held by
niche differentiation [@loreau_biodiversity_1998; @loreau_partitioning_2001] or
to the higher probability of getting high biomass species with higher species
richness [@loreau_biodiversity_2001].

@@

While we found a strong relationship between the temporal trends of community
biomass and the one of species richness, our results suggest that temporal
trends in community biomass and species richness are associated with very
different temporal trends on the connectance and average trophic level. Indeed,
temporal trends of connectance and average trophic level were positively
associated to the temporal trends of community biomass while we found
that species richness temporal trends was negatively and no
significantly related to connectance and average trophic level respectively. All
together, it suggests that temporal trends of community biomass and species
richness are related to different assembly and disassembly processes in
food-webs.

The temporal trends of connectance were positively linked to that of community
biomass, meaning that when connectance increased through time when community
biomass increased through time in a given site. Reversely, connectance decreased
in food-webs when community biomass decreased through time. Indeed, food-webs
have a highly nested structure, in a way that feeding trophic links of species
at lower trophic level are a subset of the ones of species found at higher
trophic level [@kondoh_food_2010; @nordstrom_nestedness_2015]. It is because
feeding interactions are largely driven by body size, consumers being larger
than their preys [@brose_allometric_2006; @woodward_body_2005]. We also found
that community biomass losses through time were associated to a temporal
decrease of average trophic level and that community biomass temporal gains were
associated to temporal increase of average trophic level. Together with
connectance, it suggests that community biomass losses and gains are associated
to losses and gains in large bodied piscivores, i.e. biomass at the higher
trophic level. In the same line, we found that the temporal trends of community
biomass were positively associated to the temporal trends of both number and
proportion of piscivorous fish species present in the communities (Fig.
SXX). The close association between the temporal trends of community biomass
from one side, and connectance, average trophic level from the other side,
strongly suggest that temporal trends in community biomass in natural freshwater
fish communities are mainly associated to variations in the biomass of the top
trophic level.

In contrast, we found that temporal trends of connectance was weakly negatively
related to the temporal trends of species richness, in contrasts with the strong
negative relationships that we found in the present study and from previous
results over spatial gradient [@dunne_network_2006; @danet_species_nodate].
Furthermore, we found that temporal trends of the proportion of piscivorous
species was negatively associated to the temporal trends of species richness
(Table SXX), in accordance with previous meta-analysis over spatial gradient
[@turney_pyramids_2016] and mesocosm experiment [@downing_ecosystem_2002]. It
resonates with the energetic paradigm of food-webs which proposes that the
presence of species at higher trophic level is constrained by energy loss along
the food chain [@lindeman_trophic-dynamic_1942]. It suggests that the temporal
increase of species richness found in our dataset are associated to a decrease
of connectance because species richness increases are biased toward
non-piscivorous species which have less feeding links than piscivorous species.
We found in the same time that the temporal trends species richness had no
direct link with the temporal trends of the average trophic level, the latter
being weighted by the biomass of each node. An explanation can be that when
species richness increases or decreases through time without community biomass
trends, the associated decrease and respectively increase in proportion of
piscivorous species leads to a redistribution of biomass among remaining
piscivorous species or among the non-piscivorous species.
